Inference for multivariate peaks over threshold models

### 2021-09-16 evaluation.tex
bruno's edits 2

\section{Model Comparison on the Hypersphere}
\label{sec:evaluation}
It is not immediately obvious what criteria to use to judge these models and decide which best
  represents the data's generating distribution.  As the task resides on $\mathcal{S}_{\infty}^{d-1}$,
  our options are somewhat limited in terms of what model criteria we can use.  Because of the aforementioned
  difficulty in evaluating a density directly in this space, we choose to use model evaluation
  criteria that can operate on samples from the posterior predictive distribution.  For this reason,
  we have opted to use the \emph{posterior predictive loss} criterion of \cite{gelfand1998} and the
  \emph{energy score} criterion of \cite{gneiting2007}.  Both of these metrics require calculating
  some distance in the target space, and this section will be devoted to that end.  As both these
  metrics operate on the posterior predictive distribution for each observation, we will also be
  including a metric operating on the overall posterior predictive distribution, in the form of a
  modified Kullbeck-Liebler divergence.

>>>>

\section{Scores for distributions in $\mathcal{S}_\infty^{d-1}$}
\label{sec:evaluation}

{\bf After a sentence or two motivating the need for scoring distributions in $\mathcal{S}_\infty^{d-1}$ that can be calculated with a sample-based approach, mention the paper of \cite{gneiting2007} on proper scoring rules, and introduce the energy score. Cite the result in that paper that gives conditions for the energy score to be a proper scoring rule. Then discuss the problem of finding a metric in $\mathcal{S}_\infty^{d-1}$ that corresponds to a negative definite kernel. That is, the material in Section 4.3. Introduce and discuss the idea, the write the proposition, with its proof. I don't know if the example in the 3-dimensional cube is needed or not. Finally, say how the score will be calculated in our case.
}
It is not immediately obvious what criteria to use to judge these models and decide which best
  represents the data's generating distribution.  As the task resides on $\mathcal{S}_{\infty}^{d-1}$,
  our options are somewhat limited in terms of what model criteria we can use.  Because of the aforementioned
  difficulty in evaluating a density directly in this space, we choose to use model evaluation
  criteria that can operate on samples from the posterior predictive distribution.  For this reason,
  we have opted to use the \emph{posterior predictive loss} criterion of \cite{gelfand1998} and the
  \emph{energy score} criterion of \cite{gneiting2007}.  Both of these metrics require calculating
  some distance in the target space, and this section will be devoted to that end.  As both these
  metrics operate on the posterior predictive distribution for each observation, we will also be
  including a metric operating on the overall posterior predictive distribution, in the form of a
  modified Kullbeck-Liebler divergence.

### 2021-09-16 introduction.tex
bruno's edits 2

Inferential methods for the extreme values of univariate observations
are well established and software is widely available \cite[see, for
example,][]{coles2001}. For variables in one dimension the application of EVT methods consider the asymptotic distribution of either the maxima calculated for regular blocks of data, or the values that exceed a certain threshold. The former leads to a Generalized Extreme Value (GEV) distribution, that depends on three parameters. The latter, which is the focus of this paper, leads to a Generalized Pareto (GP) distribution, that depends on a shape and a scale parameter. Likelihood-based approaches to inference can be readily implemented in both cases.
In the multivariate case the GEV theory is well established \citep[see, for example][]{dehaan2006}, but the inferential problem is more complicated, as there is no parametric representation of the GEV. This problem is inherited by the PoT approach and compunded by the fact that there is no unique definition of an exceedance of a multivariate threshold, as there is an obvious dependence on the norm that used to measure the size of a vector.

During the last decade much work has been done in the exploration of the definition and properties of an
  appropriate generalization of the univariate GP distribution 
  to a multivariate setting.  To mention some of the papers on the topic, the work of
  \cite{rootzen2006} defines the generalized Pareto distribution, with further analysis on these classes
  of distributions presented in \cite{falk2008} and \cite{michel2008}.  A recent review of the state
  of the art in multivariate peaks over threshold modelling using generalized Pareto is provided in
  \cite{rootzen2018} while \cite{RoSeWa2018a} provides insight on the theoretical properties of possible parametrizations. 
  A frequently used method for describing the dependence
  in multivariate distributions is to use a copula. \cite{renard2007,deng2011,falk2019}, provide successful examples of this approach in an EVT framework. {\bf is \cite{deng2011} really a relevant reference? It looks weird.}
  \cite{ferreira2014} presents a constructive definition of the Pareto process, that generalizes the GP to an infinite dimensional setting. It consists of decomposing the process into independent radial and angular components. Such approach can be used in the finite dimensional case, where the angular component 
  contains the information pertaining to the dependence structure of the random vector.
  Based on this definition, we present a novel approach for modelling the distribution of the angular component
  with families of distributions that provide flexibility and can be applied in a moderately large dimensional setting.
  Models for the angular measure are important for multivariate EVT applications. \cite{SaNa2014} and \cite{HaCaCh2017} consider Bayesian non-parametric approaches to model the angular measure of the multivariate GEV. Particular attention is paid to the problem of satisfying the so called moment constraint.

  The reminder of this paper is outlined as follows.

@@

For the remainder of this paper, we adopt the operators $\wedge$ to denote minima, and the $\vee$
  to denote maxima.  Thus $\wedge_i x_i = \min_i x_i$, $\vee_i x_i = \max_i x_i$.  These operators can
  also be applied component-wise between vectors, such as $\bm{a}\wedge\bm{b} = (a_1\wedge b_1, a_2\wedge b_2,\ldots)$.

\section{A model for the multivariate GP}
%\subsubsection{Univariate Maxima}
%Regarding the asymptotic behavior of extreme events, there are a couple major strategies for conducting
%  inference---developing probabilistic estimates of extreme behavior.  First developed is the theory
%  of a limiting distribution on maxima---the largest observation from a sample.  For a sample $\bm{x}$
%  where $\bm{x} = (x_1,\ldots,x_n)$ represents a sequence of $n$ independent random variables from a
%  distribution function $F$, the distribution of the maximum $M_n$ of this sequence can be derived
%  as:
%  \begin{equation*}
%    \begin{aligned}
%      \text{Pr}\left[M_n\leq z\right] &= \text{Pr}\left[X_1 \leq z, \ldots, X_n \leq z\right]\\
%        &= \text{Pr}\left[X_1\leq z\right]\times\ldots\times\text{Pr}\left[X_n\leq z\right]\\
%        &= F(z)^n.
%    \end{aligned}
%  \end{equation*}
%  In situations where $F$ is unknown, we can seek to approximate the behavior of $F^n$ as
%  $n\rightarrow\infty$.  To ensure this does not degenerate to a point mass, we need to consider a
%  standardized sequence of maxima. If this sequence stabilizes as $n$ increases, the a limiting
%  distribution exists.  More specifically, if there exists some sequence of constants $a_n > 0$, $b_n$
%  such that:
%  \begin{equation*}
%    \text{Pr}\left[\frac{M_n - b_n}{a_n} \leq z\right] \stackrel{d}{\rightarrow} G(z)
%  \end{equation*}
%  as $n\rightarrow\infty$, then $G$ is a max-stable distribution, and $F$ is in the domain of
%  attraction of that max stable distribution.  \cite{frechet1927} originates the
%  field, identifying two limiting forms of max-stable distributions, which would become known as the
%  Fr{\'e}chet and Weibull distributions. \cite{weibull1951} expands the analysis of the Weibull
%  distribution.  \cite{fisher1928}, in what is now called the \emph{Fisher-Tippett Theorem}, identifies
%  the Fr{\'e}chet and Weibull distributions, along with an as then unnamed third form, as the three
%  limiting forms of the distribution of the maxima of a sample.  \cite{gumbel1935,gumbel1942} offer
%  analysis of that third form, now known as the Gumbel distribution.  Later works, including
%  \cite{jenkinson1955} reparameterize all three forms as special cases of a single unifying form,
%  the generalized extreme value distribution, \emph{GEV}:
%  \begin{equation*}
%    \label{eqn:gev}
%    F(m \mid \mu, \sigma, \xi) = \exp\left\lbrace-\left[1 + \xi\left(\frac{x - \mu}{\sigma}\right)\right]_{}^{-1/{\xi}}\right\rbrace.
%  \end{equation*}
%  Thus distributions in the domain of attraction of an EVD (Gumbel, Fr{\'e}chet, or Weibull) will be
%  in the domain of attraction of the GEV.

@@

To develop a multivariate PoT model for extreme values consider a $d$-dimensional random vector $\bm{Z}$, whose marginal distributions have been standardized. Thus $\bm{Z} \in {\mathbb R}^d_+ = [0,+\infty]^d$, the positive $d$-dimensional cone. Following the discussion in \cite{goix2017} we assume a form of regular variation, consisting on assuming the existing of a limiting measure $\mu$, such that.
\[
    \lim\limits_{n\to\infty}n\text{Pr}\left[\frac{1}{n}\bm{Z}\geq \bm{z}\right] = \mu\left([\bm{0},\bm{z}]^c\right),
\]
where the inequality and the interval are taken componentwise. 
In this setting, $\mu$ is the asymptotic distribution of $\bm{Z}$ in extreme regions, and it is referred to as the
  \emph{exponent measure}.  $\mu$ features the homogeneity property $\mu(tA) = \frac{1}{t}\mu(A)$. This
corresponds to the second condition in Theorem 2.1 of
\cite{ferreira2014}, applied to the finite-dimensional case.
Thus, we factorize $\bm{Z} =  R\bm{V}$, where
  $R = \pnorm{\bm{Z}}{\infty} $ describes the radial component, and
  $\bm{V} = \bm{Z}/\pnorm{\bm{Z}}{\infty} \in \mathcal{S}_{\infty}^{d-1}$ the angular component of
  $\bm{Z}$,  and ${\mathbb S}_\infty^{d-1}$ denotes the positive
orthant of the unit sphere in infinite
norm.  $\mu(\cdot)$ is then factorized as
\[
    \mu\left( \left\lbrace\bm{Z} : V(\bm{Z}) \in A , R(\bm{Z}) > r\right\rbrace \right) = r^{-1}\Phi(A),
\]
  where $\Phi(\cdot)$ is the \emph{angular measure}.  The angular measure is thus independent and
  separable from the radial component.  Furthermore,
\begin{equation} \label{MGP}
    \text{Pr}\left[\bm{V} \in A \mid R > r\right]
      = \frac{r\text{Pr}\left[\bm{V} \in A, R > r\right]}{r\text{Pr}[R > r]}\hspace{0.2cm}
      \xrightarrow[r\to\infty]{~} \hspace{0.2cm} \frac{\Phi(A)}{\Phi(\mathcal{S}_{\infty}^{d-1})}.
\end{equation}
  Thus, the angular measure corresponds to the distribution of $\bm{V}$, conditional on $\bm{R}$ being large.
  As such, it controls the dependence structure of $\bm{Z}$ in the tails. To obtain an EVT model for $\bm{Z}$, we seek to establish a flexible distribution on $\bm{V} \in \mathcal{S}_{\infty}^{d-1}$.

An alternative approach to defining the multivariate GP in Equation \eqref{MGP} is provided by Theorem 6 in \cite{RoSeWa2018a}. Using the notation in that theorem, define $\bm{S} = \log(\bm{V})$. Then $\bm{S}$ is a spectral random vector, according to Definition 5. Taking $E \sim \text{Exp}(1)$, independent of $\bm{V}$, we have that $\bm{Y} = \bm{S} + E$ follows a multivariate GP, according to Theorem 6.  Then, using Proposition 4, $\bm{Z} = e^{Y}$ is also a multivariate GP. Notice that $\bm{R} = e^E$ is distributed as a standard Pareto random variable.

\subsection{Integrated Vapor Transport}
Atmospheric rivers are temporary events, where large elongated regions of high concentrations of
  water vapor are developed in the atmosphere and carry huge amounts of water potentially thousands
  of miles.  The amount of water in transit during these events can dwarf that of terrestrial rivers.
  For the targeted region, the atmospheric river can represent a significant portion of the
  precipitation the region will experience.  Thus, such events are of great interest to meteorologists,
  hydrologists, actuaries, land-use planners, and even farmers.  The ability to estimate the likely magnitude
  and distribution of such events is thus of great importance~\citep{ralph2013,ralph2018}.

One metric by which we might identify and declare atmospheric rivers is the integrated water vapor
  transport, or \emph{IVT}.  This value represents the total amount of water
  vapor being transported in an atmospheric column---that is, a column of the troposphere above a given
  area of the surface of the Earth. Raw data are measured by dropsondes---measurement devices dropped
  out of aircraft, taking measurements as they fall to the earth.~\citep{ralph2017}  A data vector
  includes an estimate of IVT at grid cell at a period in time; where a grid cell specifies the surface
  of the earth under the atmospheric column.  We have two such datasets, in differing spatial
  resolutions---the lower resolution data splits the coast of california into 8 grid cells, while the
  higher resolution does so into 47 grid cells~\citep{guan2015}.

As IVT tracks water vapor in the atmosphere, extreme values in IVT can lead to
  extreme precipitation events.  Atmospheric rivers are necessary events, in that they carry the
  bulk of precipitation that California recieves, but extreme precipitation events can be extremely
  destructive as well.  Thus, we are interested in the distribution of extreme precipitation events,
  including their depenedence structure.  This tells us about the probability of jointly extreme
  events---where IVT is extreme in two or more grid cells, as well as identifying particular
  configurations of IVT that may be anomalous.  In California, groundwater is a managed resource.
  Extreme precipitation events significantly affect groundwater, so anomalous precipitation events
  can produce anomalous configurations of groundwater, leading to destructive flooding.

We are specifically interested in extremal dependence---the relationship between the upper tails of
  dimensions of the distribution.  In this case, that means the relationship between extreme values
  in different grid cells.  We are going to be looking at point-in-time behavior rather than
  considering the time series nature of the data---that relationship may come later.

As we are interested in the extremal dependence, it makes sense that we would choose to represent
  this data using tools from extreme value theory.  Extreme value theory, or \emph{EVT}, seeks to
  model and assess probability of observing extreme events.  Such a topic is applicable generally,
  but it finds particularly strong use among such fields as finance \citep{allen2013},
  climatology \citep{trepanier2018}, and insurance \citep{beirlant1994}.  In these fields, extreme
  events may represent significant loss to the body commissioning the study.  For instance, an
  insurance company might commission a study on extreme weather events, as an extreme weather event
  localized to a particular region could cause a spike in claims from that region.  Extreme value
  theory offers us a tool set for making inference about the tails of a distribution, without having
  observed said tails.  For
  instance, with an extreme weather event like flooding, we can make predictions about return
  levels---the average time until an observation of a particular magnitude occurs---without having seen
  an observation of that magnitude, or having observed that long.  In the context of our motivating
  example, extreme values in the IVT may represent the formation of an atmospheric river, which has
  dramatic effects on precipitation and ground water.  The formation of an IVT may lead to flooding
  and other negative effects, but is also contributes necessary water for irrigation and agriculture.
  Development of statistical tools regarding these extreme events becomes necessary for making informed
  decisions.

>>>>

Inferential methods for the extreme values of univariate observations
are well established and software is widely available \cite[see, for
example,][]{coles2001}. For variables in one dimension the application of EVT methods consider the asymptotic distribution of either the maxima calculated for regular blocks of data, or the values that exceed a certain threshold. The former leads to a Generalized Extreme Value (GEV) distribution, that depends on three parameters. The latter leads to a Generalized Pareto (GP) distribution, that depends on a shape and a scale parameters. Likelihood-based approaches to inference can be readily implemented in both cases.
In the multivariate case the GEV theory is well established \citep[see, for example][]{dehaan2006}, but the inferential problem is more complicated, as there is no parametric representation of the GEV. This problem is inherited by the PoT approach and compounded by the fact that there is no unique definition of an exceedance of a multivariate threshold, as there is an obvious dependence on the norm that is used to measure the size of a vector.

During the last decade much work has been done in the exploration of the definition and properties of an
  appropriate generalization of the univariate GP distribution 
  to a multivariate setting.  To mention some of the papers on the topic, the work of
  \cite{rootzen2006} defines the generalized Pareto distribution, with further analysis on these classes
  of distributions presented in \cite{falk2008} and \cite{michel2008}.  A recent review of the state
  of the art in multivariate peaks over threshold modelling using generalized Pareto is provided in
  \cite{rootzen2018} while \cite{RoSeWa2018a} provides insight on the theoretical properties of possible parametrizations. 
  A frequently used method for describing the dependence
  in multivariate distributions is to use a copula. \cite{renard2007,deng2011,falk2019}, provide successful examples of this approach in an EVT framework. {\bf is \cite{deng2011} really a relevant reference? It looks weird. Can we say something more about the copula approach? I confess my ignorance here.}
  \cite{ferreira2014} presents a constructive definition of the Pareto process, that generalizes the GP to an infinite dimensional setting. It consists of decomposing the process into independent radial and angular components. Such approach can be used in the finite dimensional case, where the angular component 
  contains the information pertaining to the dependence structure of the random vector.
  Based on this definition, we present a novel approach for modelling  the angular component
  with families of distributions that provide flexibility and can be applied in a moderately large dimensional setting.
  Our focus on the  angular measure is similar to that in \cite{SaNa2014} and \cite{HaCaCh2017}, that consider Bayesian non-parametric approaches. Yet, our approach avoids the problem of dealing with the so called moment constraint by considering a constructive definition of the multivariate GP based on the infinity norm.

  The reminder of this paper is outlined as follows. {\bf We need a roadmap. Part of the roadmap should mention the analysis of the IVT. Basically condensing the paragraph below in a couple of sentences.}

@@

Throughout the paper, we adopt the operators $\wedge$ to denote minima, and the $\vee$
  to denote maxima.  Thus $\wedge_i x_i = \min_i x_i$, $\vee_i x_i = \max_i x_i$.  These operators can
  also be applied component-wise between vectors, such as $\bm{a}\wedge\bm{b} = (a_1\wedge b_1, a_2\wedge b_2,\ldots)$.

\section{A multivariate PoT model}
%\subsubsection{Univariate Maxima}
%Regarding the asymptotic behavior of extreme events, there are a couple major strategies for conducting
%  inference---developing probabilistic estimates of extreme behavior.  First developed is the theory
%  of a limiting distribution on maxima---the largest observation from a sample.  For a sample $\bm{x}$
%  where $\bm{x} = (x_1,\ldots,x_n)$ represents a sequence of $n$ independent random variables from a
%  distribution function $F$, the distribution of the maximum $M_n$ of this sequence can be derived
%  as:
%  \begin{equation*}
%    \begin{aligned}
%      \text{Pr}\left[M_n\leq z\right] &= \text{Pr}\left[X_1 \leq z, \ldots, X_n \leq z\right]\\
%        &= \text{Pr}\left[X_1\leq z\right]\times\ldots\times\text{Pr}\left[X_n\leq z\right]\\
%        &= F(z)^n.
%    \end{aligned}
%  \end{equation*}
%  In situations where $F$ is unknown, we can seek to approximate the behavior of $F^n$ as
%  $n\rightarrow\infty$.  To ensure this does not degenerate to a point mass, we need to consider a
%  standardized sequence of maxima. If this sequence stabilizes as $n$ increases, the a limiting
%  distribution exists.  More specifically, if there exists some sequence of constants $a_n > 0$, $b_n$
%  such that:
%  \begin{equation*}
%    \text{Pr}\left[\frac{M_n - b_n}{a_n} \leq z\right] \stackrel{d}{\rightarrow} G(z)
%  \end{equation*}
%  as $n\rightarrow\infty$, then $G$ is a max-stable distribution, and $F$ is in the domain of
%  attraction of that max stable distribution.  \cite{frechet1927} originates the
%  field, identifying two limiting forms of max-stable distributions, which would become known as the
%  Fr{\'e}chet and Weibull distributions. \cite{weibull1951} expands the analysis of the Weibull
%  distribution.  \cite{fisher1928}, in what is now called the \emph{Fisher-Tippett Theorem}, identifies
%  the Fr{\'e}chet and Weibull distributions, along with an as then unnamed third form, as the three
%  limiting forms of the distribution of the maxima of a sample.  \cite{gumbel1935,gumbel1942} offer
%  analysis of that third form, now known as the Gumbel distribution.  Later works, including
%  \cite{jenkinson1955} reparameterize all three forms as special cases of a single unifying form,
%  the generalized extreme value distribution, \emph{GEV}:
%  \begin{equation*}
%    \label{eqn:gev}
%    F(m \mid \mu, \sigma, \xi) = \exp\left\lbrace-\left[1 + \xi\left(\frac{x - \mu}{\sigma}\right)\right]_{}^{-1/{\xi}}\right\rbrace.
%  \end{equation*}
%  Thus distributions in the domain of attraction of an EVD (Gumbel, Fr{\'e}chet, or Weibull) will be
%  in the domain of attraction of the GEV.

@@

To develop a multivariate PoT model for extreme values consider a $d$-dimensional random vector $\bm{Z} \in {\mathbb R}^d_+ = [0,+\infty)^d$, the positive $d$-dimensional cone. Following the discussion in \cite{goix2017} we assume a form of regular variation, consisting on assuming the existence of a limiting measure $\mu$, such that.
\[
    \lim\limits_{n\to\infty}n\text{Pr}\left[\frac{1}{n}\bm{Z}\geq \bm{z}\right] = \mu\left([\bm{0},\bm{z}]^c\right),
\]
where the inequality and the interval are taken componentwise. 
In this setting, $\mu$ is the asymptotic distribution of $\bm{Z}$ in extreme regions, and it is referred to as the
  \emph{exponent measure}.  $\mu$ features the homogeneity property $\mu(tA) = \frac{1}{t}\mu(A)$. 
Assume that $R = \max_{i} Z_i>1$ with probability one, and that $V_i = Z_i/\max_i Z_i$ is such that $E(V_i)>0$, then
we have that the second condition in Theorem 2.1 of
\cite{ferreira2014}, applied to the finite-dimensional case, is satisfied.
Thus, we have the factorization $\bm{Z} =  R\bm{V}$, where $R$ and $\bm{V}$  are independent. Thus,
for some measure $\Phi$, $\mu(\cdot)$ is factorized as
\[
    \mu\left( \left\lbrace\bm{Z} : V \in A , R > r\right\rbrace \right) = r^{-1}\Phi(A).
\]
$R$ describes the radial component, and $\bm{V} \in \mathcal{S}_{\infty}^{d-1}$ the 
angular component of $\bm{Z}$.  Here ${\mathbb S}_\infty^{d-1}$ denotes the positive
orthant of the unit sphere in infinite norm.  $\Phi(\cdot)$ is referred to as the \emph{angular measure}.   
Moreover, $R$ is distributed as a standard Pareto random variable. Conditioning on $R$, we have that
\begin{equation} \label{MGP}
    \text{Pr}\left[\bm{V} \in A \mid R > r\right]
      = \frac{r\text{Pr}\left[\bm{V} \in A, R > r\right]}{r\text{Pr}[R > r]}\hspace{0.2cm}
      \xrightarrow[r\to\infty]{~} \hspace{0.2cm} \frac{\Phi(A)}{\Phi(\mathcal{S}_{\infty}^{d-1})}.
\end{equation}
 Thus, the angular measure corresponds to the distribution of $\bm{V}$, conditional on $R$ being large.
 As such, it controls the dependence structure of $\bm{Z}$ in the tails. In fact, following \cite{ferreira2014} the distribution function of $\bm{Z}$ is given as
\begin{equation*}
	\text{Pr}[Z_1\leq z_1, \ldots ,Z_d\leq z_d] = E\left(\bigvee_{i=1}^d
\frac{V_i}{z_i\wedge 1}\right) - E\left(\bigvee_{i=1}^d
\frac{V_i}{z_i}\right) 
\end{equation*}
where the expectation is taken with respect to $\Phi$.
In view of this, to obtain a PoT model for $\bm{Z}$, we seek 
a flexible distribution on $\bm{V} \in \mathcal{S}_{\infty}^{d-1}$.

\cite{ferreira2014} refer to the model defined in Equation \eqref{MGP} as a simple 
Pareto process indexed on a finite dimensional set. An alternative approach to 
define such distribution that provides a link to the multivariate GEV is discussed 
in Remark 1 of \cite{RoSeWa2018a}. More specifically, 
using the notation in Theorem 6 of \cite{RoSeWa2018a}, define $\bm{S} = \log(\bm{V})$. 
Then, according to Definition 5, $\bm{S}$ is a spectral random vector. Taking 
$E \sim \text{Exp}(1)$, independent of $\bm{V}$, we have that $\bm{Y} = \bm{S} + E$ 
follows a multivariate GP, according to Theorem 6.  Then, using Proposition 4, 
$\bm{Z} = e^{Y}$ is also a multivariate GP. Notice that $R = e^E$ is distributed as 
a standard Pareto random variable.

A measure that is used to characterize the strength of the dependence,
in the tail, for two random variables $W_1$ and $W_2$, with marginal
distributions $F_1$ and $F_2$ is given by
\[	\chi_{12} = \lim_{u\uparrow 1} \text{Pr}[ F_1(W_1)>u|F_2(W_2)>u]  \]
\citep{coles2001}.
$\chi_{12}$ provides information about the distribution of extremes for the variable $W_1$
given that $W_2$ is very large.  When $\chi_{12}>0$, $W_1$ and $W_2$ 
are said to
be asymptotically dependent, otherwise they are asymptotically
independent. The following result provides the asymptotic dependence coefficient for two components of $\bm{Z}$. 
\begin{prop} \label{ppchi}
Suppose that $\bm{Z} = R\bm{V}$ with $R\sim Pa(1)$,
$\text{Pr}[V_i>0] = 1$ and $E(V_i)$ exists ,
$i=1, \ldots ,d$, then
\begin{equation}
	\chi_{ij} = E\left(\frac{V_i}{E(V_i)} \wedge 
\frac{V_j}{E(V_j)}\right).
\end{equation}
\end{prop}
{\em Proof:}
Denote as $F_i$ the marginal distribution of $Z_i$. To obtain $\chi_{ij}$ we need $Pr(Z_1>z_1,Z_2>z_2)$, where $z_i =
F_i^{-1}(u) = E(V_i)/(1 - u), \;i=1,\dots, d$.
Using the fact that
$V_i>0, \forall i$ almost surely, we have that the former is equal to
\[	\text{Pr}\left[R>\frac{z_i}{V_i}\vee\frac{z_2}{V_j}\right] =
E\left(1\wedge\left(\frac{z_i}{V_i}\vee\frac{z_j}{V_j}\right)
^{-1} \right)
= E\left( \frac{V_i}{z_i}\wedge\frac{V_j}{z_j}
\right) = (1 - u)E\left(
\frac{V_i}{EV_i}\wedge\frac{V_j}{EV_j}
\right) ,
\]
where the second identity is justified by the fact that $V_i$ is
bounded and $z_i\rightarrow\infty$. The proof is completed by noting
that $Pr(F_i(Z_i)>u ) = 1 - u.\hfill\Box$

For the analysis of extreme values it is of interest to calculate the multivariate conditional survival function. The following result provides the relevant expression, as a function of the angular measure.
\begin{prop}
Assume the same conditions of Proposition \ref{ppchi}. 
Let $\alpha \subset \{1, \ldots ,p\}$ be a collections of indexes. 
Then     
\begin{equation} \label{eqn:condsurv2df}
    \text{Pr}\left[\bigcap_{l \in \alpha} Z_l > z_l \right|\left. \bigcap_{l\not\in\alpha} Z_l > z_l\right] =
      \frac{E\left(\bigwedge_{k = 1}^d 1\wedge \frac{V_k}{z_k}\right)}{E\left(\bigwedge_{k \not\in\alpha}1\wedge\frac{V_k}{z_k}\right)}.
  \end{equation}
\end{prop}  
The proof uses a similar approach to the proof of Proposition \ref{ppchi}.

%\subsection{Integrated Vapor Transport}
%Atmospheric rivers are temporary events, where large elongated regions of high concentrations of
%  water vapor are developed in the atmosphere and carry huge amounts of water potentially thousands
%  of miles.  The amount of water in transit during these events can dwarf that of terrestrial rivers.
%  For the targeted region, the atmospheric river can represent a significant portion of the
%  precipitation the region will experience.  Thus, such events are of great interest to meteorologists,
%  hydrologists, actuaries, land-use planners, and even farmers.  The ability to estimate the likely magnitude
%  and distribution of such events is thus of great importance~\citep{ralph2013,ralph2018}.

%One metric by which we might identify and declare atmospheric rivers is the integrated water vapor
%  transport, or \emph{IVT}.  This value represents the total amount of water
%  vapor being transported in an atmospheric column---that is, a column of the troposphere above a given
%  area of the surface of the Earth. Raw data are measured by dropsondes---measurement devices dropped
%  out of aircraft, taking measurements as they fall to the earth.~\citep{ralph2017}  A data vector
%  includes an estimate of IVT at grid cell at a period in time; where a grid cell specifies the surface
%  of the earth under the atmospheric column.  We have two such datasets, in differing spatial
%  resolutions---the lower resolution data splits the coast of california into 8 grid cells, while the
%  higher resolution does so into 47 grid cells~\citep{guan2015}.

%As IVT tracks water vapor in the atmosphere, extreme values in IVT can lead to
%  extreme precipitation events.  Atmospheric rivers are necessary events, in that they carry the
%  bulk of precipitation that California recieves, but extreme precipitation events can be extremely
%  destructive as well.  Thus, we are interested in the distribution of extreme precipitation events,
%  including their depenedence structure.  This tells us about the probability of jointly extreme
%  events---where IVT is extreme in two or more grid cells, as well as identifying particular
%  configurations of IVT that may be anomalous.  In California, groundwater is a managed resource.
%  Extreme precipitation events significantly affect groundwater, so anomalous precipitation events
%  can produce anomalous configurations of groundwater, leading to destructive flooding.

%We are specifically interested in extremal dependence---the relationship between the upper tails of
%  dimensions of the distribution.  In this case, that means the relationship between extreme values
%  in different grid cells.  We are going to be looking at point-in-time behavior rather than
%  considering the time series nature of the data---that relationship may come later.

%As we are interested in the extremal dependence, it makes sense that we would choose to represent
%  this data using tools from extreme value theory.  Extreme value theory, or \emph{EVT}, seeks to
%  model and assess probability of observing extreme events.  Such a topic is applicable generally,
%  but it finds particularly strong use among such fields as finance \citep{allen2013},
%  climatology \citep{trepanier2018}, and insurance \citep{beirlant1994}.  In these fields, extreme
%  events may represent significant loss to the body commissioning the study.  For instance, an
%  insurance company might commission a study on extreme weather events, as an extreme weather event
%  localized to a particular region could cause a spike in claims from that region.  Extreme value
%  theory offers us a tool set for making inference about the tails of a distribution, without having
%  observed said tails.  For
%  instance, with an extreme weather event like flooding, we can make predictions about return
%  levels---the average time until an observation of a particular magnitude occurs---without having seen
%  an observation of that magnitude, or having observed that long.  In the context of our motivating
%  example, extreme values in the IVT may represent the formation of an atmospheric river, which has
%  dramatic effects on precipitation and ground water.  The formation of an IVT may lead to flooding
%  and other negative effects, but is also contributes necessary water for irrigation and agriculture.
%  Development of statistical tools regarding these extreme events becomes necessary for making informed
%  decisions.
